ScheduleInst: Using Dyninst to Fuzz Test Multithreaded Program Schedules

### 2014-12-03 paper/paper.tex
YAY 9 PAGES

\documentclass[10pt,]{article} % For LaTeX2e
\usepackage{hyperref}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{url}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}
\bibliographystyle{unsrt}
%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09

\title{ScheduleInst: Using DynInst to Fuzz Test Multithreaded Program Schedules}
\author{Robert Wespetal\\
\and
Zachary Welch}

@@

\begin{abstract}
\setlength{\baselineskip}{18pt}

@@

subsection{Original Fuzzing}

@@

subsection{Dyninst}

@@

%\subsection{Other Multithreaded Debuggers (Liblit, etc)}

@@

	ScheduleInst is a tool designed to alter thread priorities in some user specified way, with the goal of moving away from the normal execution path.  It is our belief that testing these less probable interleavings of code that bugs will be revealed in a relatively short amount of time.  Threads are altered by using Dyninst to insert code snippets which alter the priority of a thread at a user-specified place. Dyninst is also used to provide sufficient logging to discover approximately where and how a bug manifested itself. ScheduleInst is primarily designed to find deadlocks and program crashes.  Other types of concurrency bugs like race conditions can be tested in some cases, for instance in pbzip2 we can check if any data was corrupted in the process of compressing the file, though our logging in that case would likely be unhelpful.  
	ScheduleInst has a range of options which can affect how often thread priorities are altered.  Users can specify the sequence of priorities to pass to the mutatee to be randomly generated, all priorities set to equal, or users can provide their own sequence to allow arbitrary scheduling schemes to be emulated using ScheduleInst.  Users can also specify at what points in execution priorities can be changed; by default priorities are set when a thread is created and after all thread synchronization function calls.  The granularity of logging can also be altered by the user; by default we log entrances and exits of all user-defined functions (for implementation reasons we do not instrument signal handlers) as well Pthread functions.  We allow this granularity to be changed because we have found for some applications like ffmpeg the thousands of user-defined functions generates prohibitively large execution logs for no additional information.    
	For the defualt randomly-generated thread priorities, scheduleInst allows users to specify a random seed. This feature allows users to easily recreate the environment which lead to a failed run.  Note that this is not the same thing as saying the failing runs are guaranteed to be reproducible; scheduleInst can only recreate the priorities of the threads, it cannot force the OS to yield the CPU at exactly the same point in the execution as it did in a failing run.  Additionally, we have found that in practice it is difficult to consistently  reproduce runs with identical priority sequences, and the rate seems to be dependent on the specifics of a given bug. In some cases the bug only revealed itself on 2% of runs with a known buggy priority sequence.  In other cases the bug appears regularly.  Due to the fact that consistently generating the bug is not guaranteed,  the rationale for providing thorough logging information is that we want to make it as easy as possible to find the bug with a single run.

@@

@@

At each instrumentation point, scheduleInst checks the current thread’s scheduler, sets it to the real time round robin scheduler if necessary, and alters the thread’s real-time priority to some other (usually random) valid priority value.  If the pthread\_create function is one of the instrumentation points, the new thread will be created with the round robin scheduler and appropriate priority.  Initially scheduleInst only set thread priority as part of the create call, but after testing we found setting thread priorities at the time of thread creation and after each synchronization mechanism call (eg mutex lock) stresses the application more than simply setting priorities initially and allowing the program to run.  Intuitively, our goal in altering thread priorities after synchronization mechanisms is to encourage threads to give up the cpu immediately after grabbing a lock, which would allow other threads to run and potentially enter a deadlock state if one exists.

@@

\subsection{Bugs we looking for}

@@

@@

@@

\subsection{Programs tested}

@@

@@

\subsection{Bugs found for each program}

\subsection{Types of bugs}

>>>>

\documentclass[10pt,]{article} % For LaTeX2e
\usepackage{hyperref}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{url}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{listings}
\usepackage{parcolumns}
\usepackage{subcaption}
\usepackage{multirow}
\bibliographystyle{unsrt}
%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09

\title{ScheduleInst: Using Dyninst to Fuzz Test Multithreaded Program Schedules}
\author{Robert Wespetal\\
\and
Zachary Welch}

@@

\begin{abstract}
\setlength{\baselineskip}{18pt}
ScheduleInst is a tool designed to Fuzz Test multithreaded programs.  SecheduleInst uses dynamic instrumentation to force applications to use real time threads.  By frequently altering real time thread priorities, ScheduleInst attempts to reveal concurrency bugs by generating unusual application executions. By randomly changing real time thread priorities at thread synchronization calls and with the help of dynamically instrumented logging techniques, ScheduleInst has been successful in finding real bugs in large, actively used programs. We test ScheduleInst with five popular open source applications and three simple proof of concept cases.

@@

\subsection{Original Fuzzing}

@@

\subsection{Dyninst}

@@

@@

	ScheduleInst is a tool designed to alter thread priorities in some user specified way, with the goal of moving away from the normal execution path.  It is our belief that testing these less probable interleavings of code that bugs will be revealed in a relatively short amount of time.  Threads are altered by using Dyninst to insert code snippets which alter the priority of a thread at a user-specified place. Dyninst is also used to provide sufficient logging to discover approximately where and how a bug manifested itself. ScheduleInst is primarily designed to find deadlocks and program crashes.  Other types of concurrency bugs like race conditions can be tested in some cases, for instance in pbzip2 we can check if any data was corrupted in the process of compressing the file, though our logging in that case would likely be unhelpful.  
	ScheduleInst has a range of options which can affect how often thread priorities are altered.  Users can specify the sequence of priorities to pass to the mutatee to be randomly generated, all priorities set to equal, or users can provide their own sequence to allow arbitrary scheduling schemes to be emulated using ScheduleInst.  Users can also specify at what points in execution priorities can be changed; by default priorities are set when a thread is created and after all thread synchronization function calls.  The granularity of logging can also be altered by the user; by default we log entrances and exits of all user-defined functions (for implementation reasons we do not instrument signal handlers) as well Pthread functions.  We allow this granularity to be changed because we have found for some applications like ffmpeg the thousands of user-defined functions generates prohibitively large execution logs for no additional information.    
	For the defualt randomly-generated thread priorities, scheduleInst allows users to specify a random seed. This feature allows users to easily recreate the environment which lead to a failed run.  Note that this is not the same thing as saying the failing runs are guaranteed to be reproducible; scheduleInst can only recreate the priorities of the threads, it cannot force the OS to yield the CPU at exactly the same point in the execution as it did in a failing run.  Additionally, we have found that in practice it is difficult to consistently  reproduce runs with identical priority sequences, and the rate seems to be dependent on the specifics of a given bug. In some cases the bug only revealed itself on 2\% of runs with a known buggy priority sequence.  In other cases the bug appears regularly.  Due to the fact that consistently generating the bug is not guaranteed,  the rationale for providing thorough logging information is that we want to make it as easy as possible to find the bug with a single run.

@@

ScheduleInst uses real time threads to exert some level of control over how threads run.  POSIX threads provide two real time options: SCHED\_FIFO and SCHED\_RR for First-In-First-Out and Round Robin scheduling respectively.  By default ScheduleInst uses SCHED\_RR, though in practice we found little difference between the two.  Real time schedulers run by scheduling higher priority threads first until they give up the CPU by going to sleep or ending.  In round robin the concept of time slices only exists for threads with equal priorities. The value of using real time threads for ScheduleInst is that thread priorities can be explicitly set at any point during program execution.  Unlike the default Linux scheduler (SCHED\_OTHER) which uses a multilevel feedback queue and internally alters the priority of threads based on how they use their time slices, realtime scheduling allows us to declare a certain thread “more important” to the operating system.  In real time threads, a lower priority only gets the CPU when all higher priority threads have been put to sleep.  This level of control is not perfect, but we feel using real time threads is a happy medium between doing nothing and writing our own kernel scheduler to suite our needs.  ScheduleInst forces all mutatee applications into the user-specified real time thread scheduling algorithm by using Dyninst to insert the necessary code.  By switching thread priorities often enough we hope to throw the application in question into atypical yet valid states in the hope of finding bugs.  Note that in many cases this use of real time threads and the instrumented logging will lead to decreased performance.

@@

At each instrumentation point, scheduleInst checks the current thread’s scheduler, sets it to the real time scheduler if necessary, and alters the thread’s real-time priority to some other (usually random) valid priority value.  If the pthread\_create function is one of the instrumentation points, the new thread will be created with the round robin scheduler and appropriate priority.  Initially scheduleInst only set thread priority as part of the create call, but after testing we found setting thread priorities at the time of thread creation and after each synchronization mechanism call (eg mutex lock) stresses the application more than simply setting priorities initially and allowing the program to run.  Intuitively, our goal in altering thread priorities after synchronization mechanisms is to encourage threads to give up the cpu immediately after grabbing a lock, which would allow other threads to run and potentially enter a deadlock state if one exists.

@@

\subsection{Targeted Bugs}

@@

We have designed ScheduleInst’s default behavior to specifically target these two types of errors.  In the simple deadlock example provided below, the system only deadlocks if both threads grab their first lock before either grabs their second one.  In the course of execution using the default Linux scheduler, the likelihood of this is relatively small.  Using ScheduleInst, however, the thread priority is changed immediately after a lock is grabbed.  This increases the possibility that the thread will have to give up it’s CPU before it has a chance to grab the second lock.  This gives the slower thread a better chance to catch up, and in general increases the likelihood of execution running into a deadlock.

\begin{minipage}[b]{0.45\linewidth}
\centering
\lstset{language=C,label=SliceExaple}
\begin{lstlisting}[frame=single, numbers=left, mathescape,%
  title={Deadlock Thread 1}, label=scopingExample]
...
mutex_lock(lock1);
mutex_lock(lock2);
//do some work
mutex_unlock(lock2);
mutex_unlock(lock1);
\end{lstlisting}
\end{minipage}
\hspace{0.5cm}
\begin{minipage}[b]{0.45\linewidth}
\centering
\begin{lstlisting}[frame=single,numbers=left, mathescape,%
  title={Deadlock Thread 2}, label=scopesEnteredContents]
...
mutex_lock(lock2);
mutex_lock(lock1);
//do some work
mutex_unlock(lock1);
mutex_unlock(lock2);
\end{lstlisting}
\end{minipage}

Likewise in the code below, there is the possibility for dereferencing a null pointer in threadSecond if it is run before the pointer is allocated in threadFirst.  Normally, if the thread responsible for threadFirst is created before the thread responsible for threadSecond, threadFirst will likely run before threadSecond and the code will run without crashing.  However, when using real time thread priorities, it is possible that a highly prioritized threadSecond could preempt a lowly prioritized threadFirst and cause the bug to reveal itself in a crash.

\begin{minipage}[b]{0.45\linewidth}
\centering
\lstset{language=C,label=SliceExaple}
\begin{lstlisting}[frame=single,numbers=left, mathescape,%
  title={Bad Memory Access Thread 1}, label=scopesEnteredContents]
int * myPtr = NULL;
void threadFirst(){

myPtr = malloc(sizeof(int));

\end{lstlisting}
\end{minipage}
\hspace{0.5cm}
\begin{minipage}[b]{0.45\linewidth}
\centering
\begin{lstlisting}[frame=single,numbers=left, mathescape,%
  title={Bad Memory Access Thread 2}, label=scopesEnteredContents]
void threadSecond(){

*myPtr = 4;
printf("%d\n"*myPtr);

\end{lstlisting}
\end{minipage}

In both cases ScheduleInst increases the chance that the bug will occur in execution.  We may also encourage other types of bugs like race conditions and infinite loops, but the default choices made by ScheduleInst specifically target bugs which incorrectly use synchronization mechanisms. Our logging mechanisms are also targetted at making these bugs easy to locate once they occur.  ScheduleInst may make it more likely that a program corrupts it’s data, but currently ScheduleInst would have no way to locate this error in the source code unless it generated a crash.

@@

To test ScheduleInst, we took a number of sample applications, instrumented them using ScheduleInst, and repeatedly ran each application under some load a large number of times.  The goal here is not to show that ScheduleInst finds bugs in most of it’s run’s, or that it finds many bugs in each application, the primary goal at this stage of development is to study whether ScheduleInst finds bugs faster than simply running the application many times and hoping to stumble on a bug.  Since ScheduleInst dynamically instruments logging information and scheduling controls, both of which could alter the execution of the application, we test results using only logging and both logging/schedule controls.  This should demonstrate whether the use and manipulation of real time threads have an effect on the likelihood of bugs being revealed.  These tests were run in batch; a log of all thread interactions, stdout and stderr, and a core file if it was generated were saved for each iteration.

@@

Five department Virtual Machines were used to run all tests.  The VMs all used Ubuntu 12.04; similar results were found with some initial tests using ?????.  Dyninst 8.2.1 was used as our Dyninst installation.  All test applications were compiled from source with debugging information to allow Dyninst to correctly log application specific methods. A list of applications tested and their versions can be found in Table \ref{appsTest}.  These applications were chosen because they were popular Open Source applications known to make heavy use of PThreads.  Additionally we include our original test applications as a proof of concept to show that ScheduleInst finds bugs that exist but does not find bugs in a simple application with no concurrency bugs.

\subsection{Application Loadings}
\begin{table}
\centering
\begin{tabular}{ |c|c|c| }
\hline
Application & Version &Type \\ \hline \hline
pbzip2 & Version & File Compression \\ \hline
aget   & Version & File Downloading \\ \hline
ffmpeg & Version & Video Compression \\ \hline
mysql  & Version & Database\\ \hline
apache & Version & Web Server\\ \hline
Toy ThreadSafe & N/A & Safely changes, prints global value \\ \hline
Toy Deadlock & N/A & ThreadSafe with deadlock\\ \hline
Toy Bad Mem Access &N/A& ThreadSafe with null pointer access \\ \hline
\end{tabular}
\caption{Applications Tested}
\label{appsTest}
\end{table}

For each real application tested, we had to choose some way of loading the application.  The toy applications have no external inputs and thus do not have this problem.  For each application the choice of loading is an important factor in the type of bugs we might find.  We describe our application loadings below; in all cases we feel our loadings are representative of a fairly normal use case of the application.

\begin{itemize}
  \item pbzip2: To test the parallel compressor, we generated a random 100 MB file using Linux's dd command.  pbzip2 would then attempt to zip this file using 100 threads specified on the command line. 
  \item aget: Download large file (Boost source) from department web server. 
  \item ffmpeg: BOBBY 
  \item mysql: Attach to server and run mysql provided test suite 
  \item apache: BOBBY
\end{itemize}

It is likely that different bugs would be found with different loadings, and due to our lack of knowledge about the implementation details of these applications, it is possible some of these loadings are inappropriate to provide a good test of a system.  Regardless, none of our loadings are invalid inputs to their respective programs and all work successfully with no dynamic instrumentation, so we feel that while we may not be testing these tools as well as possible, we are finding bugs in them.  In addition to our logging and full ScheduleInst runs, we tested all applications extensively with no dynamic instrumentation and never hit a crash.

@@

One important design decision is how to decide when an application has deadlocked or infinite looped.  For program crashes, we simply check for the existance of a core file after every application has reached the end of it's execution. Deadlocks, however, will run indefinitely without intervention. We decided to add a timeout to our framework that would kill the running process after 4X the run time of a normal run had elapsed. So for instance a pbzip2 run on our 100 MB file took 30 seconds in all the uninstrumented runs we performed, hence we set a timeout of 120 seconds for our instrumented pbzip2.  Once the testing was complete and certain runs were flagged as potential deadlocks, it was a simple matter of manually inspecting the thread log and suspected location of the code to decide if a deadlock had indeed occured. In practice all of the runs flagged by ScheduleInst turned out to be deadlocks.

@@

\subsection{Bugs found for each program}
\begin{table}
\centering
\begin{tabular}{ |c|c|c|c|c| }
\hline
\multirow{2}{*}{Application} & \multicolumn{2}{|c|}{Just Logging} & \multicolumn{2}{|c|}{ScheduleInst} \\ \cline{2-5}
& \# Runs & \# Bugs Found & \# Runs & \# Bugs Found \\ \hline \hline
Toy ThreadSafe & 100 &X& 100 &X \\ \hline
Toy Deadlock & 100 &X& 100 &X \\ \hline
Toy Bad Mem Access & 100 &X& 100 &X \\ \hline
pbzip2 & 100 &X& 100 &X \\ \hline
aget & 100 &X& 100 &X \\ \hline
ffmpeg & 100 &X& 100 &X \\ \hline
mysql & 100 &X& 100 &X \\ \hline
apache & 100 &X& 100 &X \\ \hline

\end{tabular}
\caption{Number of Potential Bugs Found}
\label{bugsFound}
\end{table}
\subsection{Types of bugs}

\begin{table}
\centering
\begin{tabular}{ |c|c|c|c| }
\hline
Application & \# Total Bugs & \# Seg Faults & \# Hangs \\ \hline \hline
pbzip2 & X & X & X  \\ \hline
aget   & X & X & X \\ \hline
ffmpeg & X & X & X \\ \hline
mysql  & X & X & X \\ \hline
apache & X & X & X \\ \hline
Toy ThreadSafe & X & X & X \\ \hline
Toy Deadlock   & X & X & X \\ \hline
Toy Bad Mem Access & X & X & X \\ \hline
\end{tabular}
\caption{Types of Bugs found}
\label{typeBugs}
\end{table}
